Covariate analysis of UFL-Mayo-ISB RNASeq from cerebellum

### 2016-05-06 code/Rmd/HBTRC_Array.Rmd

---
title: "Covariate analysis of MSBB microarray data"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_Array.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB Microarray')
```

@@

```{r synapse.parameters, include=FALSE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB microarrays';

thisFileName <- 'MSSM_Array.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Following brain regions were analysed for differential expression between case and control

| Brain Region                        | Synapse ID |
| ----------------------------------- | ---------- |
| middle temporal                     | syn3191101 |
| superior temporal                   | syn3191103 |
| inferior temporal                   | syn3191099 |
| precuneus superior parietal lobule  | syn3191119 |
| caudal anterior cingulate           | syn3191107 |
| frontal pole                        | syn3191095 |
| para hippocampal gyrus              | syn3191109 |

>>>>

---
title: "Covariate analysis of HBTRC microarray data"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./HBTRC_Array.Rmd",
                                 parentId ="syn5570324",
                                 entityName = "HBTRC Microarray")
```

@@

```{r synapse.parameters, include=FALSE}
# Synapse parameters
parentId = 'syn5570324';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of HBTRC microarrays';

thisFileName <- 'HBTRC_Array.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Following brain regions were analysed for differential expression

| Brain Region                        | Synapse ID |
| ----------------------------------- | ---------- |
| Cerebellum                          | syn3275756 |
| Prefrontal cortex                   | syn3275756 |
| Visual cortex                       | syn3275756 |
### Data download
#### Obtain expression matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
FOLDER_ID <- 'syn3275756';
ALL_USED_IDs <- FOLDER_ID
FOLDER_OBJ <- synGet(FOLDER_ID)

# Unzip all files
FNames = unzip(FOLDER_OBJ@filePath)

# Get expression data
brain_regions = c('cerebellum', 'visualcortex', 'prefrontalcortex')
EXPR = lapply(brain_regions, function(region, fileNames){
  ind = intersect(grep(region, fileNames),grep('gene_expression.txt', fileNames))
  expr = read.table(fileNames[ind], header=T, row.names = 1)

#   ind = intersect(grep(region, fileNames),grep('gene_expression_pvals.txt', fileNames))
#   pvals = read.table(fileNames[ind], header=T, row.names = 1)
#   pvals = pvals[rownames(expr), colnames(expr)]
#   
#   expr[pvals > 0.05] = NA
#   ind = (rowSums(is.na(expr))/dim(expr)[2]) <= 0.5
#   expr = expr[ind,]

  return(expr)
}, FNames)
names(EXPR) = brain_regions

# Get metadata
METADATA1 = fread(FNames[grep('phenotype.txt',FNames)], header=T, data.table=F)
METADATA2 = read.table(pipe("cut -f1,1,2 ./hbtrc/phenotype/phenotype_description.txt"), header=T)
METADATA = full_join(METADATA2, METADATA1)

# Arrange metadata and expression
allDATA = lapply(EXPR, function(expr, metadata){
  rownames(metadata) = metadata$trait_name
  metadata = t(metadata[,-(1:2)])

  colnames(expr) = gsub('X','',colnames(expr))
  cnames.expr = sapply(colnames(expr), function(x){str_split(x,'\\.')[[1]][1]})

  expr = expr[,names(cnames.expr)[cnames.expr %in% rownames(metadata)]]
  cnames.expr = cnames.expr[cnames.expr %in% rownames(metadata)]

  newMetadata = metadata[cnames.expr,]
  rownames(newMetadata) = colnames(expr)

  return(list(expr = expr, metadata = newMetadata))
}, METADATA)

# Seperate expression
EXPR = lapply(brain_regions, function(region, allData){
  expr = allData[[region]]$expr
  colnames(expr) = paste(colnames(expr),region,sep='.')
  expr = rownameToFirstColumn(expr, 'probeIDs')
  return(expr)
}, allDATA) %>%
  plyr::join_all(type = 'full')

# Seperate metadata
ALL.METADATA = lapply(brain_regions, function(region, allData){
  metadata = allData[[region]]$metadata
  metadata = cbind(data.frame(metadata), data.frame(donorID = rownames(metadata)))

  rownames(metadata) = paste(rownames(metadata), region, sep = '.')
  return(metadata)
}, allDATA) %>%
  rbindlist

# Store results in synapse
write.table(EXPR, file = 'rawExpression.tsv', sep = '\t', quote=F, row.names = F)
obj = File('rawExpression.tsv', name = 'Raw Expression', parentId = CODE$properties$id)
obj = synStore(obj, used = as.character(ALL_USED_IDs), executed = ThisFile, 
               activityDescription = activityDescription, activityName = activityName)

# Store results in synapse
write.table(ALL.METADATA, file = 'rawMetaData.tsv', sep = '\t', quote=F, row.names = F)
obj = File('rawMetaData.tsv', name = 'Raw Metadata', parentId = CODE$properties$id)
obj = synStore(obj, used = as.character(ALL_USED_IDs), executed = ThisFile, 
               activityDescription = activityDescription, activityName = activityName)
```

### 2016-05-06 code/Rmd/MSSM_BM_10_RNASeq.Rmd

---
title: "Covariates analysis of MSBB RNASeq data of region BM 36"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_36_RNASeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 36 RNASeq')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB BM 36 RNASeq samples';

thisFileName <- 'MSSM_BM_36_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                        | 
  | :----- | --------------------------- | 
  | BM_36  | Parahipocampal Gyrus (PHG)  |

  * BM 36: Ectorhinal area 36 is a subdivision of the cytoarchitecturally defined temporal region of cerebral cortex.

@@

METADATA = left_join(METADATA %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_TECHNICAL %>%
                       dplyr::select(BB, BM.44.Barcode, BM36.Fnl, BM22.Fnl, BM10.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_44 = BM.44.Barcode,
                                     BM_10 = BM10.Fnl,
                                     BM_22 = BM22.Fnl,
                                     BM_36 = BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  dplyr::filter(BrainRegion == 'BM_36')
rownames(METADATA) = METADATA$NewBarcode

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 36 RNASeq", parentId = parentId)
CODE <- synStore(CODE)

>>>>

---
title: "Covariates analysis of MSBB RNASeq data of region BM 10"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_10_RNASeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 10 RNASeq')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB BM 10 RNASeq samples';

thisFileName <- 'MSSM_BM_10_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                        | 
  | :----- | --------------------------- | 
  | BM_10  | Frontal Pole (FP)           |

* BM 10: frontopolar prefrontal cortex, rostrolateral prefrontal cortex, or anterior prefrontal cortex,

@@

METADATA = left_join(METADATA %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_TECHNICAL %>%
                       dplyr::select(BB, BM.44.Barcode, BM36.Fnl, BM22.Fnl, BM10.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_44 = BM.44.Barcode,
                                     BM_10 = BM10.Fnl,
                                     BM_22 = BM22.Fnl,
                                     BM_36 = BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  dplyr::filter(BrainRegion == 'BM_10')
rownames(METADATA) = METADATA$NewBarcode

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 10 RNASeq", parentId = parentId)
CODE <- synStore(CODE)

### 2016-05-06 code/Rmd/MSSM_BM_22_RNASeq.Rmd

---
title: "Covariates analysis of MSBB RNASeq data of region BM 10"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_10_RNASeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 10 RNASeq')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB BM 10 RNASeq samples';

thisFileName <- 'MSSM_BM_10_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                        | 
  | :----- | --------------------------- | 
  | BM_10  | Frontal Pole (FP)           |

* BM 10: frontopolar prefrontal cortex, rostrolateral prefrontal cortex, or anterior prefrontal cortex,

@@

METADATA = left_join(METADATA %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_TECHNICAL %>%
                       dplyr::select(BB, BM.44.Barcode, BM36.Fnl, BM22.Fnl, BM10.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_44 = BM.44.Barcode,
                                     BM_10 = BM10.Fnl,
                                     BM_22 = BM22.Fnl,
                                     BM_36 = BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  dplyr::filter(BrainRegion == 'BM_10')
rownames(METADATA) = METADATA$NewBarcode

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 10 RNASeq", parentId = parentId)
CODE <- synStore(CODE)

>>>>

---
title: "Covariates analysis of MSBB RNASeq data of region BM 22"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_22_RNASeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 22 RNASeq')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB BM 22 RNASeq samples';

thisFileName <- 'MSSM_BM_22_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                        | 
  | :----- | --------------------------- | 
  | BM_22  | Superior Temporal Gyrus     |

* BM 22: In humans it corresponds approximately to the lateral and caudal two thirds of the superior temporal gyrus,

@@

METADATA = left_join(METADATA %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_TECHNICAL %>%
                       dplyr::select(BB, BM.44.Barcode, BM36.Fnl, BM22.Fnl, BM10.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_44 = BM.44.Barcode,
                                     BM_10 = BM10.Fnl,
                                     BM_22 = BM22.Fnl,
                                     BM_36 = BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  dplyr::filter(BrainRegion == 'BM_22')
rownames(METADATA) = METADATA$NewBarcode

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 22 RNASeq", parentId = parentId)
CODE <- synStore(CODE)

### 2016-05-06 code/Rmd/Mayo_Cerebellum_RNASeq.Rmd

DE = ldply(DE, .id = 'Comparison')
tmp = DE %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(Comparison) %>%
  dplyr::summarize(nDiffExpGenes = length(unique(hgnc_symbol)),
                   DiffExpGenes = paste(unique(hgnc_symbol), collapse = ','))
kable(tmp)
```
### Store files in synapse
```{r synapse_store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum RNASeq', parentId = parentId)
CODE <- synStore(CODE)

>>>>

DE = ldply(DE, .id = 'Comparison')
tmp = DE %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(Comparison) %>%
  dplyr::summarize(nDiffExpGenes = length(unique(hgnc_symbol)))
kable(tmp)
```
### Store files in synapse
```{r synapse_store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum RNASeq', parentId = parentId)
CODE <- synStore(CODE)
